---
title: "p8105_hw2_wz2590"
output: github_document
author: Weiheng Zhang
---


```{r}
library(tidyverse)
library(dplyr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

Instacart is an online grocery service that allows people in NYC to shop online from local stores. 
The "instacart" dataset contains contains 1,384,617 observations, where each observation represents a product from an order. There are 15 variables for each observation:  
order_id: order identifier
product_id: product identifier
add_to_cart_order: order in which each product was added to cart
reordered: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
user_id: customer identifier
eval_set: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train” eval_set)
order_number: the order sequence number for this user (1=first, n=nth)
order_dow: the day of the week on which the order was placed
order_hour_of_day: the hour of the day on which the order was placed
days_since_prior_order: days since the last order, capped at 30, NA if order_number=1
product_name: name of the product
aisle_id: aisle identifier
department_id: department identifier
aisle: the name of the aisle
department: the name of the department
  
  
The key variable we will use for the following analysis is "aisle".

```{r}
data("instacart")
```

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>% 
  rename(count = n)
```

```{r fig.width=8,fig.height=8}
instacart %>% 
  count(aisle) %>% 
  rename(count = n) %>% 
  filter(count > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, count)
  ) %>% 
  ggplot(aes(x = count, y = aisle)) +
  geom_point() +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  xlab("Number of Items") + ylab("Aisles") +
  labs(title = "Top 39 Aisles Number of Items Ordered For Each")
```

This is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”, Including the number of times each item is ordered.

```{r instacart_table1}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  rename(ordered_times = n) %>% 
  mutate(rank = min_rank(desc(ordered_times))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  select(aisle, product_name, ordered_times) %>% 
  knitr::kable()
```

The following human-readable table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r instacart_table2}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```


## Problem 2
























